---
title: "Workshop on Data Visualization"
# get rid of the runtime: shiny if you want just pdf
#runtime: shiny
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R Markdown document is made interactive using Shiny. Unlike the more traditional workflow of creating static reports, you can now create documents that allow your readers to change the assumptions underlying your analysis and see the results immediately. 

To learn more, see [Interactive Documents](http://rmarkdown.rstudio.com/authoring_shiny.html).

# Motivation for data visualization


* picture of Anscombe data goes here

Load necessary packages for plotting and manipulating data:

```{r}
print("shmoeba")
print(3+3)
```


```{r include=FALSE}
library(ggplot2)
library(data.table)
source('code/dataVizUtil/R/functions.R')
```

Import the data set into a data.table:

```{r}
anscombe <- data.table(read.csv('data/anscombe.csv'))
```
What does the raw data look like? Use this command to view the "head" of the data.table.

```{r}
head(anscombe)
```


```{r}
ggplot(anscombe,aes(x=x,y=y)) + facet_grid(~set)
```


* picture of data goes here

Import your data again by first reading a file then converting it to the ***data.table*** format.
This time, we're reading a .txt file, so we can use the function 'read.table'

```{r}
dt.grd <- data.table(read.table('data/gremlinContacts_962-1345.txt',header = TRUE))

```
data.tables in R make it *very* easy to manipulate data in numerous kinds
of ways.  Here we are adjusting the numbering scheme of the amino acid residues 
by adding the integer 961 to all residue numbers .

```{r}
# fix amino acid numbering
dt.grd[,i := i+961]
dt.grd[,j := j+961]
```

'symmetrize' is a custom function I built, which adds the mirror image of data along the x = y axis. Functions are called like this:

```{r}
dt.grd.sym <- symmetrize(dt.grd)
```

Since this is a custom function, you can see what the function does by examining the code:

```{r}
symmetrize
```

Many of these lines of are data.table operations.


<!-- ## Inputs and Outputs -->

<!-- You can embed Shiny inputs and outputs in your document. Outputs are automatically updated whenever inputs change.  This demonstrates how a standard R plot can be made interactive by wrapping it in the Shiny `renderPlot` function. The `selectInput` and `sliderInput` functions create the input widgets used to drive the plot. -->

<!-- ```{r eruptions, echo=FALSE} -->
<!-- inputPanel( -->
<!--   selectInput("n_breaks", label = "Number of bins:", -->
<!--               choices = c(10, 20, 35, 50), selected = 20), -->

<!--   sliderInput("bw_adjust", label = "Bandwidth adjustment:", -->
<!--               min = 0.2, max = 2, value = 1, step = 0.2) -->
<!-- ) -->

<!-- renderPlot({ -->
<!--   hist(faithful$eruptions, probability = TRUE, breaks = as.numeric(input$n_breaks), -->
<!--        xlab = "Duration (minutes)", main = "Geyser eruption duration") -->

<!--   dens <- density(faithful$eruptions, adjust = input$bw_adjust) -->
<!--   lines(dens, col = "blue") -->
<!-- }) -->
<!-- ``` -->

<!-- ## Embedded Application -->

<!-- It's also possible to embed an entire Shiny application within an R Markdown document using the `shinyAppDir` function. This example embeds a Shiny application located in another directory: -->

<!-- ```{r tabsets, echo=FALSE} -->
<!-- shinyAppDir( -->
<!--   system.file("examples/06_tabsets", package = "shiny"), -->
<!--   options = list( -->
<!--     width = "100%", height = 550 -->
<!--   ) -->
<!-- ) -->
<!-- ``` -->

<!-- Note the use of the `height` parameter to determine how much vertical space the embedded application should occupy. -->

<!-- You can also use the `shinyApp` function to define an application inline rather then in an external directory. -->

<!-- In all of R code chunks above the `echo = FALSE` attribute is used. This is to prevent the R code within the chunk from rendering in the document alongside the Shiny components. -->



